My Findings

### 2024-03-10 Findings (Shared Folder)/Chapters/MolecularBiology.tex
new updates to biology notes

 There are two general balancing forces acting on the angiogenesis
\begin{itemize}
	\item Inhibitors:
	\begin{itemize}
		\item endostatin
		\item angiostatin
		\item thrombospondin
	\end{itemize} 
	\item Angiogens
	\begin{itemize}
		\item VEGF: Vascular Endothelial Growth Factors.
		\item bFGF: Basic Fibroblast Growth Factor.
		\item PDGF: Platelet Driven Growth Factor.
	\end{itemize}
\end{itemize}

Endothelial cells that are to form a new capillary, grow out from the side of an existing capillary by forming long pseudopodia pioneering the formation of new capillary sprout that hallows out to form a tube. This process continues until the sprout encounters another capillary, where they merge.

@@

\subsubsection*{Formation of tube structures by endothelial cells}
It was one of my main concerns that what is the process in which a single lining of endothelial cells following a tip cell forms a hallow tube (i.e. vessel). The following text from \cite{Alberts2002} explains this clearly. This process has also been described in \cite{angiogenesisYoutube}.
	\begin{wrapfigure}{r}{0.6\textwidth}
	\centering
	\includegraphics[width=0.6\textwidth]{images/hallowStructureEndothelialCells.jpg}
	\caption{The endothelial cells, when supported by suitable growth medium and signals, start to form hallow structure, that do not contain any blood, and not fluid passes through them. This indicates that the no mechanical trigger (i.e. pressure) is required to form the hallow structure for the new vessels. Image from \cite{Alberts2002}.}
\end{wrapfigure}

	``Experiments in culture show that endothelial cells in a medium containing suitable growth factors will spontaneously form capillary tubes, even if they are isolated from all other types of cells (Figure 22-26). The capillary tubes that develop do not contain blood, and nothing travels through them, indicating that blood flow and pressure are not required for the initiation of a new capillary network. Endothelial cells in culture spontaneously develop internal vacuoles that appear to join up from cell to cell, giving rise to a network of capillary tubes. These photographs show successive stages in the process.''

>>>>

\begin{observation}
	The difference in the surface properties of endothelial cells on the arterial and venous side of the developing networks of vessels control the rate an which one piece of growing capillary tube joins another. This becomes very interesting if we consider it along the observations in \cite{Koery2024}. They observed that the blunt-ended capillaries with small diameter are more susceptible for degradation after irradiation. And since the presence of blunt-ended vessels with small diameter increase the flow resistance of the network, pruning these branches ``normalizes'' the blood flow, hence increase the perfusion after irradiation.
\end{observation}

\subsubsection{Steps involved in angiogenesis}
Individual endothelial cells responds to the signals produced by the organ that they invade. The signal is complex, but the main part of the signal is vascular endothelial growth factor (\textbf{VEGF}) (which is a distant relative of platelet driven growth factor (\textbf{PDGF})). The control on the production of VEGF is through its mRNA stability and its rate of transcription. Under a low oxygen concentration, the  intracellular concentration of an active form of gene regulatory protein called \textbf{hypoxia inducible factor 1 (HIF-1)} increases. HIF-1 stimulates the transcription of VEGF gene (and the production of other genes that are needed when the oxygen supply is low). When the VEGF protein is secreted, it is then diffuses through the tissue and acts on nearby endothelial cells.

Endothelial cells that are to form a new capillary, grow out from the side of an existing capillary by forming long pseudopodia pioneering the formation of new capillary sprout that hallows out to form a tube. This process continues until the sprout encounters another capillary, where they merge. In the tumor micro environment, The growth rate of tumor increases abruptly as soon as the vessels reach it.

There are two general balancing forces acting on the angiogenesis
\begin{itemize}
	\item Inhibitors:
	\begin{itemize}
		\item endostatin
		\item angiostatin
		\item thrombospondin
	\end{itemize} 
	\item Angiogens
	\begin{itemize}
		\item VEGF: Vascular Endothelial Growth Factors.
		\item bFGF: Basic Fibroblast Growth Factor.
		\item PDGF: Platelet Driven Growth Factor.
	\end{itemize}
\end{itemize}
\FloatBarrier

\subsubsection{The Response of Endothelial Cells to VEGF}
The response of endothelial cells to VEGF has four components. First, they produce proteases to digest through the basal lamina of the parent vessels. For the second step, they migrate towards the source of VEGF, and for the third step they proliferate. Finally, they form hallow tubes. It is worth mentioning that VEGF stimulates endothelial cells selectively, while other angiogens, like fibroblast growth factor stimulates other cell types as well. The following figure summarizes these steps.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{images/endothelialResponseToVEGF.jpg}
	\caption{A summary of the response of the endothelial cells to VEGF.  Under low oxygen concentration, the intracellular concentration of HIF-1 increases. This gene regulatory protein in turn increases the transcription of VEGF protein. Then VEGF diffuses through the tissue and stimulates the endothelial cells lining the vessels. Figure is from \cite{Alberts2002}.}
\end{figure}

@@

\subsubsection*{Formation of tube structures by endothelial cells}

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{images/hallowStructureEndothelialCells.jpg}
	\caption{The endothelial cells, when supported by suitable growth medium and signals, start to form hallow structure, that do not contain any blood, and not fluid passes through them. This indicates that the no mechanical trigger (i.e. pressure) is required to form the hallow structure for the new vessels. Image from \cite{Alberts2002}.}
\end{figure}

It was one of my main concerns that what is the process in which a single lining of endothelial cells following a tip cell forms a hallow tube (i.e. vessel). The following text from \cite{Alberts2002} explains this clearly. This process has also been described in \cite{angiogenesisYoutube}.

\begin{quote}
	Experiments in culture show that endothelial cells in a medium containing suitable growth factors will spontaneously form capillary tubes, even if they are isolated from all other types of cells (Figure 22-26). The capillary tubes that develop do not contain blood, and nothing travels through them, indicating that blood flow and pressure are not required for the initiation of a new capillary network. Endothelial cells in culture spontaneously develop internal vacuoles that appear to join up from cell to cell, giving rise to a network of capillary tubes. These photographs show successive stages in the process.
\end{quote}
